Inference using context-dependent models of mutation

### 2017-05-04 writeups/context-dependence.tex
cpg results

under the model
      \begin{tabular}{c@{\quad$\to$\quad}c@{\quad at rate\quad }cc}
        $x$  &    $y$  &  $m_{xy}$ & when $x \neq y \in \{\nA,\nC,\nG,\nT\}$  \\
        \nC\nG     &  \nT\nG     &  $0.4$ & \\
        \nC\nG     &  \nC\nA     &  $0.4$ & \\
        \nT\nC\nC  &  \nT\nT\nC  &  $0.1$ & \\
        \nA\nG\nG  &  \nA\nA\nG  &  $0.1$ &
      \end{tabular} ,
with $m_{AT}=m_{TA}=0.1$, $m_{CG}=m_{GC}=0.15$, $m_{AC}=m_{TG}=m_{AG}=m_{TC}=0.08$, and $m_{CA}=m_{GT}=m_{CT}=m_{GA}=0.12$.

@@

parameter estimates differ from the truth by less than X\%.

\begin{table}
  \begin{center}
    \includegraphics[width=\textwidth]{writeup-plots/tasep_all-mcmc-runs}
  \end{center}
  \caption{ Results from iterative fitting to a $10^6$ sequence evolved to have 28.5\% sequence divergence
    with the above mutational motifs.
    \label{tab:cpg_results} }
\end{table}

>>>>

under the model given in Table \ref{tab:cpg_results},
obtained by embellishing the basic CpG hypermutability
with the longer mutational motif identified by \citet{harris}.
%     \begin{center}
%       \begin{tabular}{c@{\quad$\to$\quad}c@{\quad at rate\quad }cc}
%         $x$  &    $y$  &  $m_{xy}$ & when $x \neq y \in \{\nA,\nC,\nG,\nT\}$  \\
%         \nC\nG     &  \nT\nG     &  $0.4$ & \\
%         \nC\nG     &  \nC\nA     &  $0.4$ & \\
%         \nT\nC\nC  &  \nT\nT\nC  &  $0.1$ & \\
%         \nA\nG\nG  &  \nA\nA\nG  &  $0.1$ &
%       \end{tabular} ,
%     \end{center}
% with $m_{AT}=m_{TA}=0.1$, $m_{CG}=m_{GC}=0.15$, $m_{AC}=m_{TG}=m_{AG}=m_{TC}=0.08$, and $m_{CA}=m_{GT}=m_{CT}=m_{GA}=0.12$.

@@

parameter estimates differ from the truth by less than 3\% (mostly within 1\%).

% df <- structure(list(
%     sim = c(0.1, 0.1, 0.15, 0.15, 0.08, 0.08, 0.08, 0.08, 0.12, 0.12, 0.12, 0.12, 0.4, 0.1), 
%     fit = c(0.0996, 0.0982, 0.1505, 0.1497, 0.0784, 0.0792, 0.0798, 0.0795, 0.1202, 0.1196, 0.1198, 0.1193, 0.3978, 0.1033)), 
%     .Names = c("sim", "fit"), 
%     row.names = c("A->T", "T->A", "C->G", "G->C", "A->C", "T->G", "A->G", "T->C", "C->A", "G->T", "C->T", "G->A", "CG->TG|CG->CA", "TCC->TTC|AGG->AAG"), 
%     class = "data.frame")

\begin{table}
  \begin{center}
        \begin{tabular}{c@{\quad$\to$\quad}c@{\quad at rate\quad }ll}
          \hline
            x & y & truth & fitted \\ 
          \hline
          \nA  &   \nT           & 0.10 & 0.0996 \\ 
          \nT  &   \nA           & 0.10 & 0.0982 \\ 
          \nC  &   \nG           & 0.15 & 0.1505 \\ 
          \nG  &   \nC           & 0.15 & 0.1497 \\ 
          \nA  &   \nC           & 0.08 & 0.0784 \\ 
          \nT  &   \nG           & 0.08 & 0.0792 \\ 
          \nA  &   \nG           & 0.08 & 0.0798 \\ 
          \nT  &   \nC           & 0.08 & 0.0795 \\ 
          \nC  &   \nA           & 0.12 & 0.1202 \\ 
          \nG  &   \nT           & 0.12 & 0.1196 \\ 
          \nC  &   \nT           & 0.12 & 0.1198 \\ 
          \nG  &   \nA           & 0.12 & 0.1193 \\ 
       \nC\nG  &  \nT\nG         & 0.40 & 0.3978 \\ 
       \nC\nG  &  \nC\nA         & 0.40 & 0.3978 \\ 
    \nT\nC\nC  &  \nT\nT\nC      & 0.10 & 0.1033 \\ 
    \nA\nG\nG  &  \nA\nA\nG      & 0.10 & 0.1033 \\ 
           \hline
        \end{tabular}
  \end{center}
  \caption{ Results from iterative fitting to a $10^6$ sequence evolved to have 28.5\% sequence divergence
    with the above mutational motifs. The transitions $CG \to TG$ and $CG \to CA$ 
    were constrained to be equal (DNA strand symmetry),
    as were the transitions $TCC \to TTC$ and $AGG \to AAG$.
    \label{tab:cpg_results} }
\end{table}
